\textbf{Elimination Of Jump Statements}

### 2022-05-09 paper/main.tex
Update main.tex

\begin{abstract}
In this article we raise the problem of the so-called Jump Statements. Getting rid of the use of operators such as \textbf{goto}, \textbf{return}, \textbf{continue}, \textbf{break}, and \textbf{throw} in code is the task of most modern programming languages, as the program becomes unreadable, poorly optimized and the probability of fixing errors in such a program is extremely high.

The result of our research is an algorithm that allows converting a program in EO (Elementary Objects) language containing \textbf{GOTO} objects into a semantically similar one, but without using them.
\end{abstract}

@@

We propose an algorithm that replaces code fragments using the GOTO object with similar fragments, but without using it. Two algorithms are semantically equivalent if with the same set of input arguments they return the same results.

\section{Methods}
In EOLANG, all jump statements are implemented as a single \textbf{GOTO}-object. The main task is to distinguish what this object is in different places of the program - \textbf{goto}, \textbf{continue}, \textbf{break} or \textbf{return}. We need to learn to recognize the type of implied object and replace it with similar code accordingly.

\subsection{GOTO statement}
There are two types of \textbf{goto}-statements in the program - "forward jumps" and "backward jumps".
\subsubsection{forward jumps}
In a program "forward jump" can be used within the body of \textbf{if}-statement. In this case program will look like this:

@@

\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  if.
   not (x.eq 0)?? IS THIS SYNTAX WRONG??
   r.write (42.div x)
   TRUE
  r
\end{lstlisting}

\subsubsection{backward jumps}
Backward jump is usually used similarly to a while loop and can be easily replaced by it (considering programs that are compiled and executed).

??? INSERT IMAGE

@@

In this case we could easily replace construction using \textbf{goto} with the \textbf{while}:
\begin{lstlisting}
[] > f
 memory > i
 seq > @
  i.write 1
  i.write (i.add 1)
  while.
   i.less 10
   seq
    i.write (i.add 1)
 stdout "Finished!"
\end{lstlisting}

\subsubsection{Complex case}
??? when forward and then backward jump

\subsection{continue}

\subsection{break}

>>>>

\begin{abstract}
In this article we raise the problem of the so-called Jump Statements. Getting rid of the use of operators such as \textbf{goto}, \textbf{return}, \textbf{continue}, \textbf{break}, and \textbf{throw} in code is the task of most modern programming languages, as the program becomes unreadable, poorly optimized and the probability of fixing errors in such a program is extremely high. 
We are trying to answer the following research question: “Is it possible to completely get rid of the \textbf{GOTO} operator in the code?”

\end{abstract}

@@

The result of our research is an algorithm that allows converting a program in EO (Elementary Objects) language containing \textbf{GOTO} objects into a semantically similar one, but without using them. We propose an algorithm that replaces code fragments using the GOTO object with similar fragments, but without using it. We claim that two algorithms are semantically equivalent if with the same set of input arguments they return the same results.

\section{Methods}
In EOLANG, all jump statements are implemented as a single \textbf{GOTO}-object. The \textbf{GOTO}-object represents two types of jumps: \textit{forward jump} and \textit{backward jump}. Operators such as \textbf{break}, \textbf{return} and \textbf{goto forward} are implemented via \textit{forward jump}; \textbf{continue} and \textbf{goto backward} are implemented via \textit{backward jump}.

For each of them, a semantically identical replacement is required so that the algorithm returns the same output with the same input parameters.

\subsection{forward jump}
\textbf{\textit{Goto forward}} is implemented in the language as:
\begin{lstlisting}
+import org.eolang.gray.goto
+import org.eolang.io.stdout

goto
 [g]
  seq
   g.forward TRUE
   stdout "Will never be printed"
\end{lstlisting}

In a program \textit{forward jump} can be used within the body of \textbf{if}-statement. In this case program will look like this:

@@

Modified code:
\begin{lstlisting}
[x] > f
 memory > r
 seq > @
  r.write 0
  if.
   not. (eq. (x 0))
   r.write (42.div x)
   TRUE
  r
\end{lstlisting}

If the program encountered unconditional \textit{forward jump}, we need to transform it as \textbf{goto} with an additional condition: \textbf{if (true) goto forward}. Then the replacement occurs just as if the condition exists.

\subsubsection{return}
As mentioned above, return statement is also implemented via \textit{goto forward} object. Consider the following code:
\begin{lstlisting}
[a b] > max
 goto > @
  [g]
   seq > @
    if.
     a.greater b
     g.forward a
     TRUE
    b
\end{lstlisting}
When using goto as \textit{return} statement, an additional attribute is added to the \textbf{goto}-object, similar to the return value of a procedure. In this case, it can easily be transformed into an \textbf{if}-statement as shown in the paragraph above:
\begin{lstlisting}
[a b] > max
 if.
  a.greater b
  a
  TRUE
 b
\end{lstlisting}

\subsubsection{break}
Break is a control-flow statement which could also be implemented using \textbf{goto forward}. Since the break statement can only be found inside the generator (in EOlang Generator is a routine that can be used to control iteration behaviour of a loop) Its replacement is slightly different from the one presented in the paragraph above. In this case, we need to invert the \textit{if}-condition and add it to the condition of the loop with the logical operator \textit{AND}. After that, we easily get rid of the \textit{if}-condition itself.

Let's look at the example:
\begin{lstlisting}
[]
 memory > i
 goto > @
  [g]
   seq > @
    while.
     i.less 10
     seq
      i.write (i.add 1)
      if.
       eq. (i.mod 5 0)
       g.forward
       TRUE
\end{lstlisting}

We invert \textit{condition} and add it to \textbf{while}-condition:
\begin{lstlisting}
[]
 memory > i
 seq > @
  while.
   and.
    i.less 10
    not. (eq. (i.mod 5 0))
   seq
    i.write (i.add 1)
\end{lstlisting}

\subsection{backward jumps}
Backward jump is usually used similarly to a while loop and can be easily replaced by it (considering programs that are compiled and executed). Substitution will look like \textbf{do-while} construction.

\includegraphics{backward1}

@@

In this case we could easily replace construction using \textbf{goto} with the \textbf{while}:

\includegraphics{backward2}

\begin{lstlisting}
[] > f
 memory > i
 seq > @
  i.write 1
  i.write (i.add 1)
  while.
   i.less 10
   seq
    i.write (i.add 1)
 stdout "Finished!"
\end{lstlisting}

\subsubsection{continue}
Continue is also a control-flow statement, which occurs in loops and is implemented through the EO object \textbf{GOTO}. In order to replace it, we need to invert the \textit{if}-condition in the loop and all the statements that occur after \textbf{goto} are put in the \textit{if} body with the inverted condition.

Example of using \textit{continue} via \textbf{goto backward}:
\begin{lstlisting}
[]
 memory > i
 goto > @
  [g]
   seq > @
    while.
     i.less 10
     seq
      i.write (i.add 1)
      if.
       eq. (i.mod 5 0)
       g.backward
       TRUE
      stdout "Not divisible by 5!\n"
\end{lstlisting}

Replace the GOTO object with the described method:
\begin{lstlisting}
[]
 memory > i
 seq > @
 while.
  i.less 10
  seq
   i.write (i.add 1)
   if.
    not. (eq. (i.mod 5 0))
    stdout "Not divisible by 5!\n"
    TRUE
\end{lstlisting}

\subsection{Complex cases}
??? when forward and then backward jump
??? and mb add throw ???

\subsection{Final algorithm}
Knowing how to replace a goto-forward statement and a goto-backward statement, we can present a final algorithm to get rid of all the GOTO objects in the EOlang code.
